---
title: "Metaanalysis of COVID-19 incidence velocities"
author: 
  - "Vincent J. Carey, stvjc at channing.harvard.edu"
  - "Sean Davis, seandavi at gmail.com"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Metaanalysis of COVID-19 incidence velocities}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction
```{r doload}
suppressPackageStartupMessages({
library(sars2death)
})
```

The `Arima*` functions of sars2death model COVID-19 incidence series.
Using ARIMA(0,1,3) by default we get an estimate of drift and the
standard error of this estimate.

For the US as a whole, as of May 20 2020:
```{r lkd1, cache=FALSE}
library(sars2death)
ej = enriched_jhu_data()
usa_full = Arima_nation(ej, max_date="2020-05-20")
usa_full
```

For the US dropping NY:
```{r lkd2, cache=FALSE}
nyd = nytimes_state_data()
drny = Arima_drop_state(ej, nyd, max_date="2020-05-20")
drny

```{r lkd, fig.width=11, fig.height=5,cache=FALSE}
par(mfrow=c(1,2), mar=c(4,3,2,2))
plot(usa_full, main="USA aggregated", ylim=c(15900,36000))
plot(drny, main="USA excluding NY", ylim=c(15900,36000) )
```

Two different data sources are used for this display, as
certain anomalies were apparent on May 17 2020 that
inhibited summing NYT data to get US aggregate.  The
reconciliation of NYT, JHU and USA facts daily data could be
of interest.

# Incidence velocity estimates for contiguous 48 states and DC

```{r doall,cache=FALSE}
allst = contig_states_dc()
allarima = lapply(allst, function(x) Arima_by_state(nyd, x))
names(allarima) = allst
drifts = sapply(allarima, function(x) coef(x$fit)["drift"])
searima = function(a) sqrt(a$fit$var.coef["drift", "drift"])
se.drifts = sapply(allarima, searima)
```

```{r lkmeta}
library(rmeta)
meta.summaries(drifts, se.drifts)
```{r dorm, fig.height=10}
o = order(drifts)
metaplot(drifts[o], se.drifts[o], labels=allst[o], cex=.7, 
  xlab="Incidence velocity (CHANGE in number of confirmed cases/day)", ylab="State")
segments(rep(-350,46), seq(-49,-4), rep(-50,46), seq(-49,-4), lty=3, col="gray")
```

# Optimization of AR and MA parameters

The dependence structure of fixed-length incidence series
that are the focus of this package varies over time and between states.
The `min_bic` function will choose BIC-minimizing AR and MA orders by searching
over a grid.
We have tools to survey states in this way from a specified time point, and
to report summaries.
```{r lklkaaa}
data(min_bic_2020_05_20)
min_bic_2020_05_20
min_bic_2020_05_20[["New York"]]
```
